Are we learning to build better models at the cost of better decisions?

### 2021-03-15 manuscript/manuscript.Rmd
:pen:

Global change issues are complex and outcomes are difficult to predict, making approaches to uncertainty a central part
of effective decision-making [@ @]. Reducing that uncertainty by improving models in response to more data
becomes available

@@

<!-- 
It may be easy to dismiss this example on the argument that both models are (or more precisely,
the any logistic growth model, regardless of parameterization would be) extremely naive
and unrealistic, and any solution would surely begin with replacing both models with something
more realistic.  Yet such objections overlook the simple fact that for all its faults in 
prediction, model 1 already provides nearly optimal performance.  That performance arises
because model 1 has very accurately reflected one key aspect of the true model (the position of the
maximum growth rate), even while it is wrong about every other aspect of the biology.

An important aspect of this example is that model 1 performs nearly optimally.

@@

A wide range of paradigms are available for approaching the issue of decision-making under uncertainty. These approaches can roughly be divided into two groups: the first group treats the issue of model uncertainty independently from the decision itself, while the second integrates the process of reducing model uncertainty into the process of decision making to maximize the value of some objective. There are a wide range of techniques within each, and it is also possible to blend approaches. The key distinction is that methods in the first group do not involve any direct consideration of the possible actions or the utility that may result from those actions in how they select models [statistical approaches such as information criteria e.g. @Burnham1998]; and in particular, forecasting evaluation @Clark2001; @Dietze2018], while those in the second group require a more explicit statement of possible actions and the desired objectives [Decision theoretic approaches, for which @Polasky2011 provides an excellent and accessible review]. I illustrate how the most promising techniques from each of these groups would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcomes.  In retrospect, it will become obvious that neither these nor any other widely applied methods will select the model that leads to the best outcomes from the set of models considered.

@@

@Gneiting2007 provides a rigorous proof for such "proper" scoring criteria, which have the desirable property which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. That is, unlike likelihood or other goodness-of-fit scores, it is impossible to overfit when conditioning on a strictly proper score -- no model model can out-perform the true model. Not that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

<!-- This scoring bit comes abruptly (and it feels like could be in the introduction) -->

@@

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the simulation. Research has long emphasized the importance of learning and adaption in the face of new data whenever we are dealing with model uncertainty [@Polasky2011]. A related limitation is that in Scenario B, decisions were based either on assuming Model 1 is correct or assuming Model 2 is correct. A more robust approach [e.g. @Walters1978] incorporates the uncertainty over models directly into the decision calculations by integrating over the probability each model $M$ (or parameter value) is correct when calculating the utility; $U(x_t, a_t) = \int U(x_t, a_t | M) P(M)$. The probability assigned to each model can then be updated after each subsequent action in light of the resulting outcomes [@Ludwig1982, @Smith1981]. While a the wide range of strategies for such iterative updating are known [@Polasky2011]; all will fail to select the higher-performing of the two simple models considered  here.

@@

Decision-theoretic approaches include scenario analysis, resilience thinking, optimal control and related methods [@Polasky2011], which have a long history in ecology and particularly in fisheries management. One of the most influential of these is _adaptive management_ [sensu @Walters1978], in which a manager seeks to both reduce uncertainty over time while also achieving the best outcomes given current knowledge.  While the term today is frequently used in a looser sense, adaptive management as originally developed can be considered an example of optimal control under model uncertainty [@Walters1981; @Smith1981].  In this approach, a manager assigns probabilities to each of the possible models under consideration, i.e. models 1 & 2.  The expected utility of any action reflects both the intrinsic uncertainty over future states under either model (process uncertainty) and the manager's uncertainty over the choice of models (or parameter values). Unlike the forecast comparison, an adaptive management approach explicitly considers the utility function, Eqn \eqref{utility}, which it seeks to maximize by integrating over model uncertainty. After observing the consequences of the action, the manager updates the posterior probabilities that each model is correct, typically by Bayes rule [e.g. @Walters1981; @Smith1981; @Ludwig1982; @Punt2016]. The decision strategy is then re-evaluated in light of this revised uncertainty.  Adaptive management may be either passive or active. Under passive adaptive management, the decision-maker chooses the action which maximizes that expected future utility.  Under active adaptive management, the decision-maker may chose 'exploratory' actions which are not optimal under the current uncertainty but may lead to more rapidly decreasing the uncertainty in the future [@Walters1978]. In practice, many managers are reluctant to employ exploratory strategies, so passive adaptive management approaches such as as Management Strategy Evaluation (MSE) are more common [@Punt2016].

@@

Unfortunately, any adaptive management strategy which updates posterior distributions that characterize model undertainty leads to worse outcomes than an approach which does no updating.  I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty.  To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 4]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating [Fig 2, achieving a net present value over 100 time steps that only 31% that expected under management using Model 1 alone]. If instead of considering only these two models we consider a whole suite of Gordon-Schaefer-type models with varying $r$ and $K$ parameters, learning is slightly slower but no less counter-productive: mean stock size is $46%$ and net present value is $20%$ of what would be achieved under Model 1 (Appendix, Fig S1).

@@

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Appendix B, Fig S1).

<!-- Intuition, Figure 4 -->
The reason for Model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, Model 3.  Plotting the growth rate functions of each model, [Fig 4a], it is hardly surprising that no method exists which would not prefer the closely overlapping Model 2 to the no-where-close Model 1 as the better approximation of Model 3. However, Nevertheless, decisions based on Model 1 are nearly indistinguishable from those based on the true model [Fig 4b], while Model 2 leads to over-harvesting.  The explanation comes from noticing that the stock size corresponding to the maximum growth rate under Model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for Model 3.  Meanwhile, the peak of Model 2 occurs at a substantially lower stock size.  While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.  This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

<!-- Inttuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.  For example, Management Strategy Evaluation [@Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.  This approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.  Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem. Doing so, we would find the best constant-mortality solution under Model 1 performs much worse than the best constant-mortality solution under Model 2. Does this mean MSE is not susceptible to this issue?  No, it does not.  It is just as easy to construct an alternative Model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.  Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954]. If Model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from Model 1 and towards worse-performing models once again. Many optimization problems share this feature in which the optimal policy depends only on a subset or ratio of model parameters, such that it is usually easy to find a similar Model 1. This example also illustrates the importance of decision constraints -- a model that gives very good outcomes when we are free to vary the havest quota each year (the optimal control problem here) may give very poor results under the constraint of constant mortality, and vice versa.

```{r figure4,  fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth.  Panel B: The optimal control policy under Model 1 is nearly identical to that under the true Model 3, while the optimal policy under Model 2 supresses stock to a much lower escapement level.", fig.width=8}
# Fig 4
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

@@

<!-- Not just overfitting -->
Note that the mechanism shown here has nothing to do with the much more familiar issue of over-fitting, in which a better-fitting model will also lead to worse outcomes [@overfit].  In fact, an model which has been over-fit will gradually be rejected by either the iterative forecasting or adaptive management approaches shown here, as these approaches continually confront the models with new data to which they had not been previously fit.  The ability to avoid over-fitting is one of the greatest appeals of any iterative management strategy. But in the example here, both Model 1 and Model 2 have the same structural complexity.  The better decision performance of Model 1 is a consequence of capturing the key attribute needed for a good decision, which is perhaps a surprisingly an all-together different criteria than fit.

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood []. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ (see Appendix B, Fig S1), or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

@@

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?  It is tempting to argue that Model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.  Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [@].  If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling.  Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 4a].  Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.  It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].  This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that decision theory, such as the intuition offered by Fig 4, can help us better understand what features of a model are essential to decision outcomes and what are not.  Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.

<!-- Model free management -->
<!-- For example, a simple approach could gradually increase harvest from 0 as long as the observed recruitment increases relative to previous values.  This corresponds to a local search for the peak growth-rate from above. This strategy is optimal for the model considered here, at least under small noise, but is not robust to growth rates with local optima, nor is it optimal during a possibly-very-long transient period of the search.   -->

@@

>>>>

Global change issues are complex and outcomes are difficult to predict, making approaches to uncertainty a central part
of effective decision-making [@ @]. Reducing that uncertainty through additional data collection and evaluation of outcomes
is essential for effective management [@].

@@

<!--Gonna be two examples.  Shorten me? -->
A wide range of paradigms are available for approaching the issue of decision-making under uncertainty. These approaches can roughly be divided into two groups: statistical approaches and adaptive management approaches.  The first group, "statistical approaches," seeks to maximize some statistical metric (such as goodness-of-fit or predictive accuracy) in determining which model(s) will drive decisions. In contrast, adaptive management approaches seek to select actions which maximize the decision-maker's expected net utility over

Models are assessed by comparison to observation and the best model or models are used to inform management decisions.  This process of model evaluation may be repeated iteratively after observing the outcome of the management decision compared to the predictions under the model. This approach is implicit to most statistical methods of model estimation and assessment [e.g. information criteria, @Burnham1998]

There are a wide range of techniques within each, and it is also possible to blend approaches. The key distinction is that methods in the first group do not involve any direct consideration of the possible actions or the utility that may result from those actions in how they select models [statistical approaches such as information criteria e.g. @Burnham1998]; and in particular, forecasting evaluation @Clark2001; @Dietze2018], while those in the second group require a more explicit statement of possible actions and the desired objectives [Decision theoretic approaches, for which @Polasky2011 provides an excellent and accessible review]. I illustrate how the most promising techniques from each of these groups would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcomes.  In retrospect, it will become obvious that neither these nor any other widely applied methods will select the model that leads to the best outcomes from the set of models considered.

<!-- Proper scores-->
In the first group, I focus on evaluation using "proper" scoring rules as defined by @Gneiting2007, rather than potentially more familiar model comparisons such as $r^2$, likelihood, or information criteria [@Burnham1998] because proper scoring rules are more robust  (see @Gneiting2007 for comparison and how the approaches relate under appropriate assumptions). @Gneiting2007 provides a rigorous proof of scoring rules that have the desirable property of being "proper", that is, rules for which no model predicting the distribution of future outcomes, $Q(x)$ can achieve a better average score than the true model $P(x)$. Unlike likelihood or other goodness-of-fit scores alone, it is impossible (in expectation) to overfit the proper score -- no model model can out-perform the true model. Note that strictly proper scoring rules score _probabilistic forecasts_ and not just point predictions, favoring models which accurately reflect the uncertainty over those which under-estimate it.  These features have made proper scoring rules for probabilistic forecasts a successful and popular approach for addressing model uncertainty in many other areas [@Gneiting2014; @Raftery2016] and a promising tool for evaluation of ecological forecasts [@Dietze2018].

<!--Adaptive management-->
For the second group, decision-theoretic approaches, I focus on the example of adaptive management
Decision-theoretic approaches include scenario analysis, resilience thinking, optimal control and related methods [@Polasky2011], which have a long history in ecology and particularly in fisheries management. One of the most influential of these is _adaptive management_ [sensu @Walters1978], in which a manager seeks to both reduce uncertainty over time while also achieving the best outcomes given current knowledge.  While the term today is frequently used in a looser sense, adaptive management as originally developed can be considered an example of optimal control under model uncertainty [@Walters1981; @Smith1981].  In this approach, a manager assigns probabilities to each of the possible models under consideration, i.e. models 1 & 2.  The expected utility of any action reflects both the intrinsic uncertainty over future states under either model (process uncertainty) and the manager's uncertainty over the choice of models (or parameter values). Unlike the forecast comparison, an adaptive management approach explicitly considers the utility function, Eqn \eqref{utility}, which it seeks to maximize by integrating over model uncertainty. After observing the consequences of the action, the manager updates the posterior probabilities that each model is correct, typically by Bayes rule [e.g. @Walters1981; @Smith1981; @Ludwig1982; @Punt2016]. The decision strategy is then re-evaluated in light of this revised uncertainty.  Adaptive management may be either passive or active. Under passive adaptive management, the decision-maker chooses the action which maximizes that expected future utility.  Under active adaptive management, the decision-maker may chose 'exploratory' actions which are not optimal under the current uncertainty but may lead to more rapidly decreasing the uncertainty in the future [@Walters1978]. In practice, many managers are reluctant to employ exploratory strategies, so passive adaptive management approaches such as as Management Strategy Evaluation (MSE) are more common [@Punt2016].

@@

@@

@@

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the simulation. It would be possible to generate forecasts for the next $n$ steps, but these forecasts would also be conditional on the management action (i.e. harvest quota) selected at each step.  To evaluate two-step-ahead predictions we must consider each possible action under each possible state predicted by the step-ahead forecast, weighted by probability of the state given the model and the probability of the model.  This rapidly expanding set of possibilities is addressed by adaptive management for sequential decision problems [e.g. @Smith1981], which I employ in the next section.

@@

@@

Any adaptive management strategy updates posterior distributions over model uncertainty [@Punt2016; @Ludwig1982]. Unfortunately, any such adaptive updating leads to worse outcomes than the equivalent non-adaptive strategy, in which model uncertainty is held fixed.  I illustrate the application of a passive adaptive management strategy to this simple example, following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty. Passive adaptive management for a simple sequential decision problem is straight forward to implement over a discrete set of states and actions using dynamic programming with iterative updates [@Smith1981, example code in Appendix]. To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig 3]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating [Fig 2, achieving a net present value over 100 time steps that only 31% that expected under management using Model 1 alone].

@@

Increasing the space of possible models to cover a whole plausible range of parameters $r$ and $K$ does nothing to resolve the problem [Fig 4].

@@

<!-- quick summary of what was demonstrated -->
Given this simple decision problem in which one of the two models leads to effectively optimal ecological and economic outcomes, current approaches invariably choose the other. Moreover, a decision maker employing an adaptive management or iterative forecasting assessment such as those considered here would have no way of realizing that the outcomes they experienced were in fact sub-optimal.  In both approaches, the manager quickly concludes that Model 1 is entirely implausible, while finding that Model 2 is remarkably accurate at predicting future values. This problem is not addressed by re-estimating parameters (equivalently, considering a larger suite of models with all possible $K$ and $r$ values): in fact leads to a model with slightly lower $K$ and even higher level of over-fishing and under-performance than Model 2 (Fig 4).

<!-- Intuition, Figure 5 -->
The reason for Model 1's seemingly contradictory ability to make good decisions but bad forecasts becomes obvious once we compare both curves to that of the underlying model, Model 3.  Plotting the growth rate functions of each model, [Fig 5A], it is hardly surprising that no method exists which would not prefer the closely overlapping Model 2 to the no-where-close Model 1 as the better approximation of Model 3. However, Nevertheless, decisions based on Model 1 are nearly indistinguishable from those based on the true model [Fig 5B], while Model 2 leads to over-harvesting.  The explanation comes from noticing that the stock size corresponding to the maximum growth rate under Model 1 (the peak of the curve) falls at almost exactly the same stock size as that of the peak growth rate for Model 3.  Meanwhile, the peak of Model 2 occurs at a substantially lower stock size.  While the optimal control solution appears to depend only on a step-ahead forecast accuracy (indeed, the SDP solution method used here takes only step-ahead forecast probabilities as input, [@Marescot2013]), mathematical analysis showed long ago [@Reed1979] that the optimal solution for this problem depends only on keeping the biomass at the value responsible for the maximum growth rate.  This realization is quite general: for most decision problems, simple models will exist under which the optimal decision is the same as it would be for the true model, even when that simple model is wrong in most other ways.

<!-- Intuition for it's not just this example -->
This phenomenon is not unique to sequential decision problems or optimal control solutions.  For example, Management Strategy Evaluation [@Punt2016] in fisheries seeks to evaluate pre-specified strategies rather solve for the best possible strategy through optimal control.  This approach is well justified -- optimal control techniques such as stochastic dynamic programming illustrated here quickly become intractable for more complex models [@Marescot2013] typically used in fisheries stock assessments.  Constraining the search to only strategies that impose a constant mortality (defined as harvest per unit biomass, $F = H/B$) means we only have to evaluate those $N$ strategies each iteration, not solve a sequential decision problem. Doing so, we would find the best constant-mortality solution under Model 1 performs much worse than the best constant-mortality solution under Model 2. Does this mean MSE is not susceptible to this issue?  No, it does not.  It is just as easy to construct an alternative Model 1 with the same properties of leading to nearly optimal decisions while being rejected by any method of iterative learning, such as forecasting or adaptive management.  Just as the optimal control policy for a Gordon-Schaefer (logistic) model depends only on parameter $K$ [@Reed1979], the optimum constant escapement policy depends only on parameter $r$ [@Schaefer1954]. If Model 1 has a value of $r$ such that it happens to match the best possible constant-mortality solution for the (unobserved) true model, then $K$ can be set arbitrarily high, ensuring the any model selection, forecasting, or adaptive management approach would lead away from Model 1 and towards worse-performing models once again. Many optimization problems share this feature in which the optimal policy depends only on a subset or ratio of model parameters, such that it is usually easy to find a similar Model 1. This example also illustrates the importance of decision constraints -- a model that gives very good outcomes when we are free to vary the havest quota each year (the optimal control problem here) may give very poor results under the constraint of constant mortality, and vice versa.

```{r figure5,  fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth.  Panel B: The optimal control policy under Model 1 is nearly identical to that under the true Model 3, while the optimal policy under Model 2 supresses stock to a much lower escapement level.", fig.width=8}
# Fig 5
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

@@

<!-- Not just overfitting -->
Note that the mechanism shown here has nothing to do with the much more familiar issue of over-fitting, in which a better-fitting model will also lead to worse outcomes [@overfit].  In fact, an model which has been over-fit will gradually be rejected by either the iterative forecasting or adaptive management approaches shown here, as these approaches continually confront the models with new data to which they had not been previously fit.  The ability to avoid over-fitting is one of the greatest appeals of any iterative management strategy. In the example here, both Model 1 and Model 2 have the same structural complexity.  The better decision performance of Model 1 is a consequence of capturing the key attribute needed for a good decision, which is perhaps a surprisingly an all-together different criteria than fit. Note that other methods of model selection not considered here, including goodness-of-fit metrics such as $r^2$  or information criteria [@Burnham], will all prefer Model 2 over Model 1 for the same reason.

<!-- Not just "get more models" -->
This case cannot be dismissed merely as being dealt a bad hand in having to pick only between Model 1 and Model 2. Model 1 approximates the key feature, giving nearly optimal outcomes.  Previous literature has underscored the importance of Knightian "unknown unknowns" two alternative models examined here fail to fully reflect our uncertainty in the underlying dynamics, and the consequences of underestimating model uncertainty are well understood []. Had we included the true model in the set of possibilities, the techniques illustrated would have had little difficulty in distinguishing it from Models 1 and 2 after sufficient iterations. In practice, we never have the true model, and even a much more thorough suite of candidate models, such as considering any possible values of $r$, $K$ [Fig 4], or even among common alternative models with different structural forms, such as Beverton-Holt [@Beverton1957], Ricker [@Ricker1954], or Shepherd [@Shepherd1980], would fail to do any better, since all these models still have symmetric growth rates with a peak growth rate at half the steady-state population size).  Models will always be simpler than reality.  To insist that this issue can and will be avoided by always including more and better models of the process misses the point.

@@

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?  It is tempting to argue that Model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.  Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [@].  If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling.  Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 5A].  Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.  It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].  This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that  the intuition offered by decision theory [Fig 5] can help us better understand what features of a model are essential to decision outcomes and what are not.  Conversely, models which appear to lead to accurate predictions (like Model 2) can result in outcomes that are far from optimal. Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.

<!-- Model free management: what would it look like? -->
<!-- For example, a simple approach could gradually increase harvest from 0 as long as the observed recruitment increases relative to previous values.  This corresponds to a local search for the peak growth-rate from above. This strategy is optimal for the model considered here, at least under small noise, but is not robust to growth rates with local optima, nor is it optimal during a possibly-very-long transient period of the search.   -->

<!-- Correlary: Neither accurate forecasting nor effective outcomes are sufficient grounds to assert model validity. (meh, that's just 'all models are wrong, some are useful' for certain things)   -->

## Acknowledgements

The author acknowledges support from NSF CAREER Award #1942280 and helpful discussions with Melissa Chapman and Jeremy Fox.

@@

<!--
